### 2022-10-12 isabelle/Doc/Specification/document/json.tex
Configure literate haskell for the json spec

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{module}\;\Conid{\Conid{Language}.\Conid{Marlowe}.Json}\;\mathbf{where}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{testLiterate}\mathbin{::}\Conid{IO}\;(){}\<[E]%
\\
\>[B]{}\Varid{testLiterate}\mathrel{=}\Varid{putStrLn}\;\text{\ttfamily \char34 Yes,~this~is~literate~haskell.~Thanks~sed!\char34}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks