### 2022-10-13 isabelle/Doc/Specification/document/json.tex
Moved json instances to isabelle

The Json specification for Marlowe Core is defined in Literate Haskell using the Aeson library. In order
to fully understand the specification, some knowledge of the library is recommended. For each type we
define a way to parse the JSON value to the datatype (FromJSON instances) and a way to serialize the
datatype to JSON (ToJSON instances).

\isamarkupsubsection{Party}

\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{9}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\mathbf{instance}\;\Conid{ToJSON}\;\Conid{Party}\;\mathbf{where}{}\<[E]%
\\
\>[B]{}\hsindent{5}{}\<[5]%
\>[5]{}\Varid{toJSON}\;(\Conid{Address}\;\Varid{address})\mathrel{=}{}\<[E]%
\\
\>[5]{}\hsindent{4}{}\<[9]%
\>[9]{}\Varid{object}\;[\mskip1.5mu \text{\ttfamily \char34 address\char34}\mathbin{.=}\Varid{address}\mskip1.5mu]{}\<[E]%
\\
\>[B]{}\hsindent{5}{}\<[5]%
\>[5]{}\Varid{toJSON}\;(\Conid{Role}\;\Varid{name})\mathrel{=}{}\<[E]%
\\
\>[5]{}\hsindent{4}{}\<[9]%
\>[9]{}\Varid{object}\;[\mskip1.5mu \text{\ttfamily \char34 role\char95 token\char34}\mathbin{.=}\Varid{name}\mskip1.5mu]{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{addressExample}\mathbin{::}\Conid{Party}{}\<[E]%
\\
\>[B]{}\Varid{addressExample}\mathrel{=}\Conid{Address}\;\text{\ttfamily \char34 example~address\char34}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{roleExample}\mathbin{::}\Conid{Party}{}\<[E]%
\\
\>[B]{}\Varid{roleExample}\mathrel{=}\Conid{Role}\;\text{\ttfamily \char34 example~role\char34}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
\par

The example address is serialized as \ensuremath{\{\mskip1.5mu \text{\ttfamily \char34 address\char34}\mathbin{:}\text{\ttfamily \char34 example~address\char34}\mskip1.5mu\}}, and the role
example as \ensuremath{\{\mskip1.5mu \text{\ttfamily \char34 role\char95 token\char34}\mathbin{:}\text{\ttfamily \char34 example~role\char34}\mskip1.5mu\}}